A Formal Model of Checked C

### 2023-01-04 sandp23/evaluation.tex

% \review{While I found the idea behind section V very interesting, the current version
%   of this section lacks some details that would help in better understanding (1) 
%   how the approach works, and (2) the overall scope of the approach. 
%   $\\$
%   For instance, the authors state that, following [19], they try to "exercise
%   interesting patterns" by adding "admissible but redundant typing rules" like
%   G-ASTR. There are a few points that are unclear here: (1) are these rules
%   discovered manually or automatically (starting from the Redex semantics)?, (2)
%   are there any guiding principles for coming up with rules that lead to
%   interesting cases?
%   $\\$
%   Later, the authors refer to "generation rules modified to be slightly more
%   permissive" to generate "a little" ill-typed terms. Again, are these rules
%   obtained automatically or defined manually? If the latter, did you follow any
%   methodology to derive such rules? Are these rules the same as the "admissible
%   but redundant typing rules" from above?}
% \liyi{Deena? Leo? }
Our evaluation of \systemname consists of a series of tests that can be categorized into Micro-benchmarks.
MicroBenchmarking involves evaluating performance on fundamental operations involving tainted pointers, context switching between checked and sandboxed regions, and sandboxed execution of functions. We further go on to evaluate \systemname on six real-world programs pertaining to diversified domains to evaluate real-world run-time and memory performance. We use C's \<time.h\> library to evaluate the Runtime performance by placing each of the test-bench calls within the timing scope (Program scope within which the timer runs). Our timing scope excludes marshaling activities within the test-case environment with an intuition that marshaling is irrelevant if tainted\-ness of a pointer is propagated from the call site up until its declaration. We use Valgrind's "massif" memory profiler to benchmark the peak memory usage of the Heap. Unlike the Runtime performance, we record Peak Memory usage as a relative offset to original program because most programs are extremely small in comparison to the constant overhead from the sandbox (81 KiB approx). This can further be optimized away by choosing to only compile custom Tainted wrappers for those STDLib functions that are in use by the tainted pointers in the program. Valgrind's memory figures do not account for Sandboxed allocations which happen in the shadow memory. 
All of the evaluation was performed using 6-Core Intel i7-10700H with 40 GB of RAM, running Ubuntu 20.04.3 LTS and the benchmarks for every test were sampled as the mean of ten consecutive iterations.

@@

\subsubsection{Pointer Access}
Pointer overhead during run-time between the pointers in the checked and Sandboxed regions is evaluated by benchmarking a test that involves 100k read/write/arithmetic operations on the pointer. 34\% in overhead is the result of "Offset To Pointer" conversion and sanity checks for Taintedness inserted by \systemname at every access to the tainted pointer.

@@

\subsection{Program Run-time Benchmarks}
\myparagraph{Overview}
We evaluate \systemname on the basis of conversion efforts and performance on six programs with the use of a pre-included test suite for each program. We intend to demonstrate \systemname's capability of enforcing spatial safety in all of the scenarios leading to modern memory vulnerabilities by re-introducing corrected memory bugs and then annotating the relevant buggy code with tainted pointers. However, this requires the bugs to be predicted before they are discovered, which is chronologically impossible. Consequently, we also annotate large sections of relevant bug-free code to mimic the developer's intuition on making \systemname annotations without any input on the bug locality. The extent of conversion qualitatively dictates the converted program's conversion efforts required and the run-time performance. Consequently, each of the conversions for the six programs follows a varied approach.

@@

\myparagraph{ProFTPD}
\systemname changes for ProFTPD are further limited in extent and aimed at the exact changes required to encapsulate CVE-2010-4221. We mark the user input to the unsafe function "pr\_netio\_telnet\_gets()" as tainted (\_TPtr$l$char$g$) and propagate its tainted-ness to its callers and callees enclosed within its defined scope. Our changes for the above function were policed by 24 unique API tests, which we use to benchmark the performance. For each test, our benchmark samples the delta between the call and return time of pr\_netio\_telnet\_gets(). This sampling is repeated 10 times and its mean value is reflected in the below table. Although \systemname is shown to require 50\% more time in servicing this request, this delta is made insignificant during deployment where network bandwidth relatively accounts for a bigger metric to the overall performance.

@@

\myparagraph{Tiny-bignum}
Motivated by its size and simplicity, we intended to make comprehensive \systemname changes for Tiny-bignum by marking all the pointers as tainted. Furthermore, in likes of an unsafe use of sprintf() that could possibly lead to a buffer overflow within the bignum\_to\_string(), we move this function into the WASM sandbox. Ideally, a major portion of conversion efforts to \systemname is attributed to understanding the codebase and finding the precise extent to which we choose to propagate the taintedness or to stop and give up and marshall the data, thereby, requiring only 4 hours to convert tiny-bignum to \systemname. The evaluation was performed on Tiny-bignum's test suite consisting of 4 Test cases, each of which test the functionality to scale on big numbers subject to all of the supported unary and binary operations.

>>>>

% \review{While I found the idea behind section V very interesting, the current version
%   of this section lacks some details that would help in better understanding (1) 
%   how the approach works, and (2) the overall scope of the approach. 
%   $\\$
%   For instance, the authors state that, following [19], they try to "exercise
%   interesting patterns" by adding "admissible but redundant typing rules" like
%   G-ASTR. There are a few points that are unclear here: (1) are these rules
%   discovered manually or automatically (starting from the Redex semantics)?, (2)
%   are there any guiding principles for coming up with rules that lead to
%   interesting cases?
%   $\\$
%   Later, the authors refer to "generation rules modified to be slightly more
%   permissive" to generate "a little" ill-typed terms. Again, are these rules
%   obtained automatically or defined manually? If the latter, did you follow any
%   methodology to derive such rules? Are these rules the same as the "admissible
%   but redundant typing rules" from above?}
% \liyi{Deena? Leo? }
Our evaluation of \systemname consists of a series of tests that can be categorized into Micro-benchmarks and Program Benchmarks.
MicroBenchmarking involves evaluating performance on fundamental operations involving tainted pointers, context switching between checked and sandboxed regions, and sandboxed execution of functions.
We further go on to evaluate \systemname on six real-world programs pertaining to diversified domains to evaluate real-world run-time and memory performance.

\myparagraph{Benchmarking Methodology}
We use C's \<time.h\> library to evaluate the Runtime performance by placing each of the test-bench calls within the timing scope (Program scope within which the timer runs). Our timing scope excludes marshaling activities within the test-case environment with an intuition that marshaling is irrelevant if tainted\-ness of a pointer is propagated from the call site up until its declaration. We use Valgrind's "massif" memory profiler to benchmark the peak memory usage of the Heap. Unlike the Runtime performance, we record Peak Memory usage as a relative offset to original program because most programs are extremely small in comparison to the constant overhead from the sandbox (81 KiB approx). This can further be optimized away by choosing to only compile custom Tainted wrappers for those STDLib functions that are in use by the tainted pointers in the program. Valgrind's memory figures do not account for Sandboxed allocations which happen in the shadow memory. 
All of the evaluation was performed using 6-Core Intel i7-10700H with 40 GB of RAM, running Ubuntu 20.04.3 LTS and the benchmarks for every test were sampled as the mean of ten consecutive iterations.

@@

\subsubsection{Pointer Access}
Run-time Pointer overhead is evaluated by benchmarking a test that involves 100k read/write/arithmetic operations on the pointer. 34\% in overhead is the result of "Offset To Pointer" conversion and sanity checks for Taintedness inserted by \systemname at every access to the tainted pointer.

@@

\subsection{Program Run-time Benchmarks}
\myparagraph{Overview}
We evaluate \systemname on the basis of conversion efforts and performance on six programs with the use of a pre-included test suite for each program. We intend to demonstrate \systemname's capability of enforcing spatial safety in all of the scenarios leading to modern memory vulnerabilities by re-introducing corrected memory bugs and then performing \systemname annotations on the relevant buggy code . However, this requires the bugs to be predicted before they are discovered, which is chronologically impossible. Consequently, we also annotate large sections of relevant bug-free code to mimic the developer's intuition on making \systemname annotations without any input on the bug locality. Conversion style and extent qualitatively dictate the converted program's conversion efforts required and the run-time performance. Consequently, each of the conversions for the six programs follows a varied approach.

@@

\myparagraph{ProFTPD}
\systemname changes for ProFTPD are further limited in extent and aimed at the exact changes required to encapsulate CVE-2010-4221. We mark the user input to the unsafe function "pr\_netio\_telnet\_gets()" as tainted (\_TPtr<char>) and propagate its tainted-ness to its callers and callees enclosed within its defined scope. Our changes for the above function were policed by 24 unique API tests, which we use to benchmark the performance. For each test, our benchmark samples the delta between the call and return time of pr\_netio\_telnet\_gets(). This sampling is repeated 10 times and its mean value is reflected in the below table. Although \systemname is shown to require 50\% more time in servicing this request, this delta is made insignificant during deployment where network bandwidth relatively accounts for a bigger metric to the overall performance.

@@

\myparagraph{Tiny-bignum}
Due to its small size and simplicity, \systemname changes for Tiny-bignum was chosen to be comprehensive. Furthermore, bignum\_to\_string() was moved to the sandbox due to a memory-unsafe use of sprintf(). Given that significant \systemname conversion efforts is attributed to understanding the source-code and finding the precise extent to which we choose to propagate the taintedness or to stop and give up to marshalling the data between regions, Tiny-bignum only required 4 hours. The evaluation was performed on Tiny-bignum's test suite consisting of 4 Test cases, each of which test the functionality to scale on big numbers subject to all of the supported unary and binary operations.

### 2023-01-04 sandp23/overview.tex

\Cref{fig:overview} provides the compiler structure of \systemname. The \systemname type checker compiles a \systemname program by partitioning the different program regions, i.e., $c$ and $uc$ regions, as well as different pointers, i.e., checked, unchecked and tainted pointers. Any $uc$ region pointers are executed within a sandbox, while checked pointers are executed via the \checkedc compiler.

\Cref{sec:formal} provides the formalism of the different compiler components.
Here,
we use the example of a simple network server in~\lst{lst:comb} to explain the \systemname compiler.
The~\lst{lst:comb} shows the C code of the redacted version of a simple network server with~\code{server_loop} as its entry point (\entrypoint).
The server runs in a loop and calls~\code{handle_request}, which handles each network request.
The function~\code{handle_request} reads data from the socket through~\code{read_msg} and based on the first byte, either~\code{process_req1} or~\code{process_req2} will be called to handle the request.

@@

\liyi{The example looks good. But we are focusing on converting C to \checkedc here? }

\myparagraph{Goal} 
The developer's goal is to partition the code in~\lst{lst:comb} so that spatial memory vulnerabilities do not affect the overall program.
Ideally, the developer must convert the entire code to Checked C so that the entire program achieves spatial memory safety.
However, Checked C has restrictions that can result in a considerable conversion effort.
For instance,~\code{void*} pointers are not directly supported by Checked C.
Consequently, the developer needs to convert functions using~\code{void*} pointers into generic versions -- this could be tedious.
To handle this, the developer can do a best-effort conversion and annotate only a few pointers~\eg by using an automated conversion tool such as~\threec{} which annotates few pointers as shown in~\lst{lst:initialconv}.
However, as explained in~\sect{subsec:nosafetyagsintuncheckedcode}, the unchecked code can also affect the safety of checked pointers.

>>>>

\Cref{fig:overview} provides the compiler structure of \systemname. The \systemname type checker compiles a \systemname program by partitioning the different program regions, i.e., $c$ and $uc$ regions, as well as different pointers, i.e., checked, unchecked and tainted pointers. Any $uc$ region pointers are executed within a sandbox, while checked pointers are executed via the \checkedc compiler within the checked region.

\Cref{sec:formal} provides the formalism of the different compiler components.
Converting a C/checked-c program to \systemname is a two-step process that involves identifying unsafe/unchecked regions described in \Cref{subsec:identifyregionstosbx} followed by making the \systemname changes described in \Cref{subsec:moveregionstosbx}.  
\subsection{Identify Unchecked Regions}
\label{subsec:identifyregionstosbx}
Here,
we use the example of a simple network server in~\lst{lst:comb} to explain changes required by the \systemname compiler.
The~\lst{lst:comb} shows the C code of the redacted version of a simple network server with~\code{server_loop} as its entry point (\entrypoint).
The server runs in a loop and calls~\code{handle_request}, which handles each network request.
The function~\code{handle_request} reads data from the socket through~\code{read_msg} and based on the first byte, either~\code{process_req1} or~\code{process_req2} will be called to handle the request.

@@

\liyi{The example looks good. But we are focusing on converting C to \checkedc here? }
\subsection{\systemname Annotations}
\label{subsec:moveregionstosbx}
\myparagraph{Goal} 
The developer's goal is to partition the code in~\lst{lst:comb} so that spatial memory vulnerabilities do not affect the overall program.
Ideally, the developer must convert the entire code to Checked C so that the entire program achieves spatial memory safety.
However, Checked C has restrictions that can result in a considerable conversion effort.
For instance,~\code{void*} pointers are not directly supported by Checked C.
Consequently, the developer needs to convert functions using~\code{void*} pointers into generic versions -- this could be tedious.
To handle this, the developer can do a best-effort conversion and annotate only a few pointers~\eg by using an automated conversion tool such as~\threec{} which annotates few pointers as shown in~\lst{lst:initialconv}.
However, as explained in~\sect{subsec:nosafetyagsintuncheckedcode}, the unchecked code can also affect the safety of checked pointers.
